---
output:
  word_document: default
  html_document: default
---

#1(a)
$$X_{1},X_{2},...,X{n}\mid Z_{i}～Z_{i}Po(\lambda_{d})+(1-Z_{i})Po(\lambda_{c})$$
$$f(X_{i})=Z_{i}P(X_{i};\lambda_{d})+(1-Z_{i})P(X_{i};\lambda_{c})$$
$$L(\lambda_{d},\lambda_{c},X,Z)=\prod_{i=1}^{n}(P(X_{i};\lambda_{d})^{Z_{i}}*P(X_{i};\lambda_{c})^{1-Z_{i}}*\pi^{Z_{i}}*(1-\pi)^{(1-Z_{i})})$$
$$\ell(\lambda_{d},\lambda_{c},X,Z)=\sum_{i=1}^{n}(Z_{i}*[log(P(X_{i};\lambda_{d}))+log(\pi)]+(1-Z_{i})*[log(P(X_{i};\lambda_{c}))+log(1-\pi)])$$
To define Q in E-step, we need $E_{Z\mid X;\pi^{(t)};\lambda^{(t)}_{d};\lambda^{(t)}_{c}}(Z_{i})$
$$\alpha^{(t)}_{i}=E_{Z\mid X;\pi^{(t)};\lambda^{(t)}_{d};\lambda^{(t)}_{c}}(Z_{i})$$
$$=P(Z_{i}=1\mid X;\lambda^{(t)}_{d};\lambda^{(t)}_{c})$$
$$=\frac{P(Z_{i}=1;X;\lambda^{(t)}_{d};\lambda^{(t)}_{c})}{P(X;\lambda^{(t)}_{d};\lambda^{(t)}_{c})}$$
$$=\frac{\sum_{Z_{i} \in\lbrace0,1\rbrace,j\neq i}P(Z_{j};Z_{i}=1;X;\lambda^{(t)}_{d};\lambda^{(t)}_{c})}{\sum_{Z_{i}\in\lbrace0,1\rbrace}P(Z_{j};X;\lambda^{(t)}_{d};\lambda^{(t)}_{c})}$$
$$=\frac{\sum_{Z_{i} \in\lbrace0,1\rbrace,j\neq i}(\prod_{k=1,k\neq 1}P(X_{k}\mid Z_{k};\lambda_{d}^{(t)};\lambda_{c}^{(t)}))P(Z_{k};\lambda^{(t)}_{d};\lambda^{(t)}_{c})P(X_{i}\mid Z_{i}=1;\lambda_{d}^{(t)};\lambda_{c}^{(t)}))P(Z_{i}=1;\lambda^{(t)}_{d};\lambda^{(t)}_{c})}{\sum_{Z_{i}\in\lbrace0,1\rbrace,j\neq i}(\prod_{k=1,k\neq1}P(X_{k}\mid Z_{k};\lambda_{d}^{(t)};\lambda_{c}^{(t)}))P( Z_{k};\lambda^{(t)}_{d};\lambda^{(t)}_{c})(\sum_{Z_{i}\in\lbrace0,1\rbrace}P(X_{i}\mid Z_{i};\lambda_{d}^{(t)};\lambda_{c}^{(t)})P(Z_{i};\lambda^{(t)}_{d};\lambda^{(t)}_{c}))}$$
$$=\frac{P(X_{i}\mid Z_{i}=1;\lambda^{(t)}_{d};\lambda^{(t)}_{c})P(Z_{i}=1;\lambda^{(t)}_{d};\lambda^{(t)}_{c})}{\sum_{Z_{i}\in\lbrace0,1\rbrace} P(X_{i}\mid Z_{i};\lambda^{(t)}_{d};\lambda^{(t)}_{c})P(Z_{i};\lambda^{(t)}_{d};\lambda^{(t)}_{c})}$$
So we can get 
$$\alpha_{i}^{(t)}=\frac{\pi^{(t)}P(X_{i};\lambda^{(t)}_{d})}{\pi^{(t)}P(X_{i};\lambda^{(t)}_{d})+(1-\pi^{(t)})P(X_{i};\lambda^{(t)}_{c})}$$
$$Q(\lambda^{(t)}_{d};\lambda^{(t)}_{c};\lambda_{d};\lambda_{c})=E_{Z\mid X;\lambda^{(t)}_{d};\lambda^{(t)}_{c}}(\ell(\lambda^{(t)}_{d};\lambda^{(t)}_{c};X,Z))$$
$$=\sum_{i=1}^n \{\alpha^{(t)}_i*[log[P(X_i;\lambda_d)]+log\pi]+(1-\alpha^{(t)}_i)*[log[P(X_i;\lambda_c)]+log(1-\pi)] \}$$
#1(b)
$$\frac{\partial Q}{\partial \pi}=\sum^n_{i=1}(\frac{\alpha^{(t)}_i}{\pi}-\frac{1-\alpha^{(t)}_i}{1-\pi})$$
$$\sum^n_{i=1}[\frac{\alpha^{(t)}_i}{\pi^{(t+1)}}-\frac{1-\alpha^{(t)}_i}{1-\pi^{(t+1)}}]=0$$
then we can get$$\pi^{(t+1)}=\frac{\sum^n_{i=1}\alpha^{(t)}_i}{n}$$
because $$log[P(X_i;\lambda)]=X_ilog(\lambda)-\lambda-log(X_i !)$$
$$\frac{\partial Q}{\partial \lambda_d}=\sum^n_{i=1}[\alpha^{(t)}_i(-1+\frac{X_i}{\lambda_d})]$$
then we can get$$\lambda^{(t+1)}_d=\frac{\sum^n_{i=1}\alpha^{(t)}_iX_i}{\sum^n_{i=1}\alpha^{(t)}_i} $$
$$\frac{\partial Q}{\partial \lambda_c}=\sum^n_{i=1}[(1-\alpha^{(t)}_i)(-1+\frac{X_i}{\lambda_c})]$$
then we can get$$\lambda^{(t+1)}_c=\frac{\sum^n_{i=1}(1-\alpha^{(t)}_i)X_i}{\sum^n_{i=1}(1-\alpha^{(t)}_i)} $$

```{r}
#1(c)
x<-t(c(2,0,0,1,3,0,1,6,2,0,1,0,2,0,8,0,1,3,2,0))
dim(x)
xmean<-mean(x)
em<-list(pai=c(0.5),lambdad=c(3.5),lambdac=c(1/3))
# compute log(1 + exp(x))
log1pe <- local({
  LOG.EPS2 <- log(.Machine$double.eps / 2)
  function (x) {
    l <- 0
    if (x > 0) { l <- x; x <- -x } # range reduction
    if (x < LOG.EPS2) l else l + log(1 + exp(x))
  }
})

# E-step auxiliary function: compute log(alpha_t)
log.alpha <- function (x, em) {
  f1 <- log(dpois(x, em$lambdad))
  f2 <- log(dpois(x, em$lambdac))
  ll <- log(em$pai / (1 - em$pai)) # logit(lambda)
  -log1pe(-ll + f2 - f1)
}
# `x` has observations in columns
# `em` stores parameters: list(lambdad;lambdac;pai)
em.update <- function (x, em) {
  # [ E-step ]
  p <- exp(apply(x, 2, log.alpha, em))
  # [ M-step ]
  pai <-mean(p)
  lambdad <- apply(x, 1, weighted.mean, p)
  lambdac <- apply(x, 1, weighted.mean, 1 - p)
  list(pai = pai, lambdad = lambdad, lambdac = lambdac)
}
em.update(x,em)

em.update.iterate<-function(x,em){
  em<-em
  em.new<-em.update(x,em)
  while(abs((em.new$pai-em$pai)/em.new$pai) >= 1e-8){
    em<-em.new
    em.new<-em.update(x,em)
  }
  em.new
}
em.update.iterate(x,em)

```

```{r}
#1(d)
emn<-em.update.iterate(x,em)
exp(log.alpha(2,emn))
exp(log.alpha(3,emn))
exp(apply(x,2,log.alpha,emn))

```
    From the result above, we can get the probalbility of the first intersetion being danger is 0.0323, and the probalbility of the fifth intersection being danger is 0.1754. By calculating all intersection, wa can think that the 5th, 8th, 15th and 18th are black spots.
```{r}
swapem<-list(pai=c(0.5),lambdad=c(1/3),lambdac=c(3.5))
em.update.iterate(x,swapem)
```

#2(a)
$$X_{j}\mid Z～N(\mu_{1}I(j\leq Z)+\mu_{2}I(j>Z),\sigma^2)$$

$$\ell(\theta,X,Z=i)=-\frac{n}{2}*(log\sigma^2)-(2\sigma^2)^{-1}\sum^{n-1}_{i=1}\lbrace I(Z=i)(X-\mu_{(i)})^T(X-\mu_{(i)})\rbrace$$
$$\ell(\theta,X_{j},Z)=-\frac{n}{2}*(log\sigma^2)-(2\sigma^2)^{-1}\sum^{n}_{j=1}\lbrace I(j\leq Z)(X_{j}-\mu_{1})^2+I(j>Z)(X_{j}-\mu_{2})^2)\rbrace$$
$$Q(\theta;\theta^{(t)})=E_{Z\mid X;\theta^{(t)}}[\ell(\theta^{(t)};X,Z)]$$
$$=-\frac{n}{2}log(\sigma^2)-\frac{1}{2\sigma^2}\sum^{n-1}_{i=1}E_{Z\mid X;\theta^{(t)}}I(Z=i)(X-\mu_{(i)})^T(X-\mu_{(i)})$$
$$=-\frac{n}{2}log(\sigma^2)-\frac{1}{2\sigma^2}\sum^{n}_{j=1}\lbrace E_{Z\mid X;\theta^{(t)}}[I(j\leq Z)](X_{j}-\mu_{1})^2+E_{Z\mid X;\theta^{(t)}}[I(j>Z)](X_{j}-\mu_{2})^2\rbrace$$
#2(b)
$$\pi_{i}^{(t)}=E_{Z\mid X;\theta^{(t)}}I(Z=i)$$
$$=P(I(Z=i)=1\mid X,\theta^{(t)})$$
$$=\frac{P(I(Z=i)=1, X,\theta^{(t)})}{P(X,\theta^{(t)})}$$
$$=\frac{\sum_{I(Z=i)\in\lbrace0,1\rbrace,k\neq i}P(I(Z=k);I(Z=i)=1;X;\theta^{(t)})}{\sum_{I(Z=i)\in\lbrace0,1\rbrace,k}P(I(Z=i)=1;X;\theta^{(t)})}$$
$$=\frac{P(X_{j}\mid I(Z=i)=1;\theta^{(t)})P(I(Z=i)=1,\theta^{(t)})}{\sum_{I(Z=i)\in\lbrace0,1\rbrace,k} P(X_{j}\mid I(Z=i)=1;\theta^{(t)})P(I(Z=i)=1;\theta^{(t)})}$$
$$=\frac{\frac{1}{n-1}\frac{1}{\sqrt{2\pi\sigma^2}}exp(-(X-\mu_{(i)}^{(t)})^T(X-\mu_{(i)}^{(t)})/2\sigma^{2(t)})}{\frac{1}{n-1}\sum_{k=1}^{n-1}\frac{1}{\sqrt{2\pi\sigma^2}}exp(-(X-\mu_{(k)}^{(t)})^T(X-\mu_{(k)}^{(t)})/2\sigma^{2(t)})}$$
$$\frac{exp(-(X-\mu_{(i)}^{(t)})^T(X-\mu_{(i)}^{(t)})/2\sigma^{2(t)})}{\sum_{k=1}^{n-1}exp(-(X-\mu_{(k)}^{(t)})^T(X-\mu_{(k)}^{(t)})/2\sigma^{2(t)})}$$
#2(c)
$$\frac{\partial Q}{\partial \mu_{1}}=\frac{1}{\sigma^2}\sum^{n}_{j=1}E_{Z\mid X;\theta^{(t)}}[I(j\leq Z)](X_{j}-\mu_{1})$$
$$\sum^{n}_{j=1}E_{Z\mid X;\theta^{(t)}}[I(j\leq Z)]X_{j}=\sum^{n}_{j=1}E_{Z\mid X;\theta^{(t)}}[I(j\leq Z)]\mu^{(t+1)}_{1}$$
so we can get $$\mu^{(t+1)}_{1}=\frac{\sum^{n}_{j=1}E_{Z\mid X;\theta^{(t)}}[I(j\leq Z)]X_{j}}{\sum^{n}_{j=1}E_{Z\mid X;\theta^{(t)}}[I(j\leq Z)]}$$
$$\frac{\partial Q}{\partial \mu_{2}}=\frac{1}{\sigma^2}\sum^{n}_{j=1}E_{Z\mid X;\theta^{(t)}}[I(j> Z)](X_{j}-\mu_{2})$$
$$\sum^{n}_{j=1}E_{Z\mid X;\theta^{(t)}}[I(j> Z)]X_{j}=\sum^{n}_{j=1}E_{Z\mid X;\theta^{(t)}}[I(j\leq Z)]\mu^{(t+1)}_{2}$$
so we can get $$\mu^{(t+1)}_{2}=\frac{\sum^{n}_{j=1}E_{Z\mid X;\theta^{(t)}}[I(j> Z)]X_{j}}{\sum^{n}_{j=1}E_{Z\mid X;\theta^{(t)}}[I(j> Z)]}$$
$$\frac{\partial Q}{\partial \sigma^2}=\frac{-n}{2\sigma^2}+\frac{1}{2(\sigma^2)^2}\sum^{n-1}_{i=1}E_{Z\mid X;\theta^{(t)}}[I(Z=i)](X-\mu_{(i)})^T(X-\mu_{(i)})$$
$$\frac{n}{2(\sigma^2)^{(t+1)}}=\frac{1}{2[(\sigma^2)^{(t+1)}]^2}\sum^{n-1}_{i=1}\pi_{i}^{(t)}(X-\mu_{(i)}^{(t+1)})^T(X-\mu_{(i)}^{(t+1)})$$
so we can get$$\sigma^{2(t+1)}=\frac{1}{n}\sum^{n-1}_{i=1}\pi_{i}^{(t)}(X-\mu_{(i)}^{(t+1)})^T(X-\mu_{(i)}^{(t+1)})$$

```{r}
data("Nile")
mu1<-mean(Nile[1:50])
mu2<-mean(Nile[51:100])
sigma<-(sum((Nile[1:50]-mu1)^2)+sum((Nile[51:100]-mu2)^2))/100


# E-step auxiliary function: compute log(pi_t)

pii<-function(x,mu1,mu2,sigma){
  nominator<-numeric(length=99)
  for (k in 1:99) {
  nominator[k]<-exp(-crossprod(x-c(rep(mu1,k),rep(mu2,100-k)))/(2*sigma))
  }
  denominator<-sum(nominator)
  nominator/denominator
}

logpii<-function(x,mu1,mu2,sig){#log(pit),sigma^2 canceled
  pt<-numeric(length = 99)
  for(i in 1:99){
    pt[i]<-(-crossprod(x-c(rep(mu1,i),rep(mu2,100-i))))
  }
  pt<-pt-log1pe(sum(pt[2:99])/pt[1])
  pt
  }

Qtheta<-function(x,mu1,mu2,sigma){
  Q<-numeric(99)
  for (i in 1:99) {
    Q[i]<-logpii(x,mu1,mu2,sigma)[i]*crossprod(x-c(rep(mu1,i),rep(mu2,100-i)))
  }
  Qtheta<--50*log(sigma)-sum(Q)/(2*sigma)
  return(Qtheta)
}


  
em2.update<-function(x,mu1,mu2,sigma){
  Q.old<-0
  Q<-Qtheta(x,mu1,mu2,sigma)
  u1<-numeric(100)
  d1<-numeric(100)
  u2<-numeric(100)
  d2<-numeric(100)
  s<-numeric(99)
  p<-numeric(99)
  while (abs(Q-Q.old)/Q>=10e-6) {
    Q.old<-Q
    for (j in 1:100){
      u1[j]<-sum(pii(x,mu1,mu2,sigma)[j:99],na.rm=T)*x[j]
      d1[j]<-sum(pii(x,mu1,mu2,sigma)[j:99],na.rm=T)
      u2[j]<-sum(pii(x,mu1,mu2,sigma)[1:j-1],na.rm=T)*x[j]
      d2[j]<-sum(pii(x,mu1,mu2,sigma)[1:j-1],na.rm=T)
    }
      mu1<-sum(u1)/sum(d1)
      mu2<-sum(u2)/sum(d2)
    for (i in 1:99) {
         s[i]<-pii(x,mu1,mu2,sigma)[i]*crossprod(x-c(rep(mu1,i),rep(mu2,100-i)))
    }
      sigma<-sum(s)/100
      p<-logpii(x,mu1,mu2,sigma)
      Q<-Qtheta(x,mu1,mu2,sigma)
  }
  list(mu1 = mu1, mu2 = mu2,sigma = sigma,pit=p)
}
int<-em2.update(Nile,mu1,mu2,sigma)
int
pit<-int$pit
plot(pit,type="l") #plot pi find the change point location
which.max(pit)  #the 28th
year<-seq(1871,1970,1)
mu28<-c(rep(int$mu1,28),rep(int$mu2,100-28))
tempdata<-data.frame(year,Nile,mu28)

```


